Accounting for uncertainty in simulating greenhouse gas emissions from dairy cattle in Kenya

### 2022-06-13 Manuscript_figures.Rmd
more progress

```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4), fig.cap = "Total farm emissions and emissions intensity for farm number 5 from the farm survey, based on the outputs of a Monte Carlo simulation. The blue bar indicates a point estimate based on the default values recommended by the IPCC"}

@@

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "Total greenhouse gas emissions per farm and emissions intensity across all farms in the farm survey, based on the outputs of a Monte Carlo simulation."}

@@

@@

Here's the relationship between these two variables. We exclude farms with a per capita milk yield of <500 kg per year, since such farms can't be considered dairy farms and they would create outliers in many analysis.

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="Relationship between milk yield per capita and emissions intensity of milk production for farms in Kenya, based on a Monte Carlo simulation. Outputs are shown as a scatter plot (left) and a density surface (right)."}
normal_milk_yield<-which(GHG_simulation_scenarios$y$pc_milk_yield>500)

@@

```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="Probability distributions of greenhouse gas emissions intensity at specific levels of milk production per capita. Slices through the emissions intensity density surface are shown for a production level of 2000 kg milk per capita and year (left) and as a comparison between production levels of 2000 and 3000 kg milk per capita and year (right)."}
slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=2000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)
slice_3000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=3000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)

@@

```{r, fig.dim = c(8, 4), fig.cap="Probability that farms that differ in per capita milk production also differ in greenhouse gas emissions intensity. The surface, which is based on the outputs of a Monte Carlo simulation, expresses the probability that a farm with higher milk productivity (y-axis) has a lower emissions intensity than a farm with lower milk productivity (x-axis)."}

@@

```{r defining_interventions}
# breed intervention

cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]

ggplot(cows, aes(factor(breed),milk_yield)) +
  geom_violin(draw_quantiles=c(0.5),fill="light green") +
  xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()

ggsave("figures/Fig_S1_breeds.png", width=7,height=5)

>>>>

```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4), fig.cap = "***Figure 1.*** *Estimated total emissions from livestock (a) and emissions intensity of milk production (b) for a smallholder dairy farm in Kenya. Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions. The blue line indicates a precise estimate based on precise results from a farm survey and default parameters recommended by various guidance documents.*"}

@@

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure 2.*** *Total greenhouse gas emissions from livestock (a) and emissions intensity of milk production (b) across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms). Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}

@@

The

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure S1.*** *Greenhouse gas emissions from livestock per head of cattle across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms). Plausible estimates are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
all_farms_per_capita<-ggplot(data=GHG_simulation_scenarios$y, aes(pc_on_farm)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") + theme_bw() +
    xlab(expression(Total~greenhouse~gas~emissions~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("Emissions per capita")

all_farms_per_capita

ggsave("figures/Fig_S1_emissions_from_all_farms_per_capita.png",width=9,height=5)

We can now look at summary statistics of total and per capita emissions from cattle across livestock-keeping smallholder farms in Kenya. To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emissions intensity, we also look at summary statistics across only the farms with per capita milk yields of >500 kg of milk.

normal_milk_yield<-which(GHG_simulation_scenarios$y$pc_milk_yield>500)

emissions_summary<-rbind(data.frame("Emissions per farm"=round(quantile(GHG_simulation_scenarios$y$on_farm,
                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions per capita"=round(quantile(GHG_simulation_scenarios$y$pc_on_farm,
                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions intensity"=round(quantile(GHG_simulation_scenarios$y$CO2eq_per_milk,
                                                                   c(0,0.05,0.5,0.95,1)),1)),
                         data.frame("Emissions per farm"=round(quantile(dairy_households$y$on_farm,
                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions per capita"=round(quantile(dairy_households$y$pc_on_farm,
                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions intensity"=round(quantile(dairy_households$y$CO2eq_per_milk,
                                                                   c(0,0.05,0.5,0.95,1)),1)))
emissions_summary$Quantile<-c("Minimum","5% quantile", "Median", "95% quantile", "Maximum",
                              "Minimum","5% quantile", "Median", "95% quantile", "Maximum")
rownames(emissions_summary)<-NULL

kable(emissions_summary[,c(4,1,2,3)], booktabs = TRUE,
      col.names = c("Distribution attribute",
                    paste0("CO","$_2$","-eq farm","$^{-1}$"),
                    paste0("CO","$_2$","-eq head","$^{-1}$"),
                    paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S1.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
  index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per capita" = 1, "Emissions intensity" = 1))

\

@@

Here's the relationship between these two variables, for farms with per-capita milk yields of >500 kg per year.

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per capita (x-axis) and emissions intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b). The vertical lines in (b) illustrate slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^ (the corresponding distributions are shown in Fig. 4b).*"}

@@

```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emissions intensity of milk production at specific levels of per-capita milk production among livestock-keeping smallholder farms in Kenya. Distributions are shown for (a) a production level of 2000 kg milk head^-1^ year^-1^, including a histogram of the random samples the distribution is based on, and (b) production levels of 2000 and 3000 kg milk head^-1^ year^-1^.*"}
slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=2000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)
slice_3000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=3000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)

@@

```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-capita milk production levels. Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-capita milk production and emissions intensity. This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}

@@

```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Observed distributions of milk yields for all cattle breeds encountered among dairy cows in a survey of livestock-keeping smallholder farms in Kenya.*"}
# breed intervention

cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
cowcount<-data.frame(table(cows$breed))

breed_yield_plot<-ggplot(cows, aes(factor(breed),milk_yield)) +
  geom_violin(draw_quantiles=c(0.5),fill="light green") +
  xlab("Breed") + ylab(expression(Milk~yield~(kg~cow^{-1}~day^{-1}))) +
  geom_boxplot(width=0.1,fill="forestgreen") +
  theme_bw() +
  coord_flip() +
  scale_x_discrete(labels=c("Holstein Friesian (pure/improved)",
                            "Ayrshire (pure/improved)",
                            "Jersey (pure/improved)",
                            "Guernsey (pure/improved)",
                            "Cross-bred unknown",
                            "Boran",
                            "Local zebu")) +
  geom_text(data=cowcount ,aes(x = Var1, y = -2, label=paste("n =",Freq)),size = 4)

breed_milk_summary<-data.frame(Breed=c("Holstein Friesian (pure/improved)",
                            "Ayrshire (pure/improved)",
                            "Jersey (pure/improved)",
                            "Guernsey (pure/improved)",
                            "Cross-bred unknown",
                            "Boran",
                            "Local zebu"),
                            cow_number=cowcount$Freq,
                            milk_yield=round(ggplot_build(breed_yield_plot)$data[[2]]$middle,1),
                            intervention_factor= round(ggplot_build(breed_yield_plot)$data[[2]]$middle/
                             ggplot_build(breed_yield_plot)$data[[2]]$middle[1],2))
breed_milk_summary$intervention_factor[which(breed_milk_summary$Breed=="Boran")]<-1

breed_yield_plot

ggsave("figures/Fig_S2_breeds.png", width=9,height=4)

```
```{r}
kable(breed_milk_summary, col.names = c("Breed", "Number of dairy cows", "Daily milk yield (kg)", "Intervention factor"),
      caption = "<p style='color:black'><i><b>Table S2.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
column_spec(2:4, width = "5cm")

\
